---
title: "Coral_Larvae_DIN_Model"
author: "Madison Enda"
format: html
editor: visual
---

## Modeling the Potential Relationship Between Dissolved Inorganic Nitrogen and Coral Larvae

```{r, echo= FALSE, message= FALSE, warning=FALSE}
# Load the required packages
library(tidyverse)
library(dplyr)
library(kableExtra)
library(here)
library(lubridate)
```

### Read in our data with appropriate variable names using here(read.csv())

```{r}
# Read in the data as applicable variables
demography <- read.csv(here("data", "MCR_LTER_Lenihan_mapped_corals_20220413.csv"))

coral_recruits <- read.csv(here("data", "coral_recruit_tile_spat_counts_2006-2016_20180626.csv"))

yearly_temp <- read.csv(here("data", "MCR_LTER_FOR04_PhysoData_Yearly_average.csv"))

algae <- read.csv(here("data", "MCR_LTER_Macroalgal_CHN_2005_to_2022_20230713.csv"))

```

### Filter our demography, algae data to only be Site 4 (where our temperature data is from)

```{r}
# Filter to only Site 4 and drop NAs
site4_demography <- demography %>%
  filter(Site == "LTER4") %>%
  drop_na()

#Selecting the algae and N data, filtering to Site 4, and dropping NAs
algae_select <- algae %>%
  filter(Site== "LTER 4") %>%
  select(Year, Dry_Weight, N) %>%
  drop_na()

```

### Filter the SST data to our desired time period (2013-2016)

```{r}
# Selecting only temperature and year from the water column data
yearly_temp_upper <- yearly_temp  %>%
  select(upper_watercolumn_temperature_sbe39_C, time_local) %>%
  drop_na()

# Filtering the years of the algae data from 2013-2019
temp_3_years <- yearly_temp_upper %>%
  filter(time_local >= 2013 & time_local < 2017) %>%
  mutate(Year= time_local) %>%
    mutate(Year = year(Year)) %>%
  mutate(Temp = upper_watercolumn_temperature_sbe39_C)%>%
  select(Year, Temp)

kable(temp_3_years)
```

### Filter the algae data and group by year to get algae and Nitrogen annual averages

```{r}
# Filtering to the years we have all data on
algae_3_years <-algae_select %>%
  filter(Year >= 2013 & Year <= 2016)

# Taking the mean nitrogen content by year
annual_average_N <- algae_3_years %>%
  group_by(Year) %>%
  summarize(mean(N))

# Taking the mean algae biomass by year
annual_average_algae <- algae_3_years %>%
 group_by(Year) %>%
  summarize(mean(Dry_Weight))
 
kable(annual_average_N)
kable(annual_average_algae)

```

### Filter our larval recruits data to years of interest and take annual mean total

```{r}
# Group by year and take the mean of adult observations
average_recruits <- coral_recruits %>%
 group_by(nominal_year) %>%
  mutate(annual_recruits= mean(sum(count)))

# Filter to 2013 through 2016
recruits_3_years <- average_recruits %>%
  filter(nominal_year >= 2013 & nominal_year <2017)

average_recruits_3_years <- recruits_3_years %>%
  group_by(nominal_year) %>%
  summarize(mean(annual_recruits))

kable(average_recruits_3_years)
```

```{r}
average_larval_recruits <- mean(average_recruits_3_years$`mean(annual_recruits)`)
average_larval_recruits
```

### Repeating the process for average recruits, but this time grouping by species as well

```{r}
# Filter for desired years, group by family and year
average_recruits_species <- coral_recruits %>%
  filter(nominal_year >= 2013 & nominal_year <= 2016) %>%
  group_by(nominal_year, family) %>%
  mutate(average_recruits = sum(count))

# Find the average number of recruits each year for each coral species
species_recruits <- average_recruits_species %>%
  group_by( nominal_year, family) %>%
  filter(family == c('Acroporidae')) %>%
  summarize(mean(average_recruits))

kable(species_recruits)
```

### Repeat the demography filtering process to get information for each species

```{r}
# Group by taxa and year, then take the average number of observations
adult_population <- site4_demography %>%
 group_by(Year, Taxa) %>%
  mutate(annual_pop= mean(n()))

# Group by taxa to find the average number of adults for each species
species_population <- adult_population %>%
  group_by( Year, Taxa) %>%
  summarize(mean(annual_pop))

kable(species_population)
```

### Creating a data frame with the data for each species, to create models for Pocilloporidae, Acroporidae, and Poritidae groups.

```{r}
acro_data <- data.frame(Year = annual_average_algae$Year, 
                        Algae = annual_average_algae$`mean(Dry_Weight)`, 
                        Nitrogen = annual_average_N$`mean(N)`,
                        Temp = temp_3_years$Temp,
                        Adult_Pop = c(9, 33, 39, 38),
                        Larval_Recruits = c(38, 31, 51, 27))


poci_data <- data.frame(Year = annual_average_algae$Year, 
                        Algae = annual_average_algae$`mean(Dry_Weight)`, 
                        Nitrogen = annual_average_N$`mean(N)`,
                        Temp = temp_3_years$Temp,
                        Adult_Pop = c(99, 63, 63, 64),
                        Larval_Recruits = c(197, 74, 58, 261))



pori_data <- data.frame(Year = annual_average_algae$Year, 
                        Algae = annual_average_algae$`mean(Dry_Weight)`, 
                        Nitrogen = annual_average_N$`mean(N)`,
                        Temp = temp_3_years$Temp,
                        Adult_Pop = c(109, 141, 114, 73),
                        Larval_Recruits = c(29, 29, 55, 41))
```

```{r}
kable(acro_data) %>%
  kable_classic_2(full_width = F)
```


### Predicting Coral Cover Instead of Adult Population

```{r}
# Loading in the coral cover data
coral_cover <- read.csv(here("data", "knb-lter-mcr.4_1_20240105.csv"))

#Selecting the date, taxa, and % cover data, filtering to Site 4, and dropping NAs
coral_cover_site4 <- coral_cover %>%
  filter(Site== "LTER 4") %>%
  select(Date, Taxonomy_Substrate_or_Functional_Group, Percent_Cover) %>%
  drop_na()


# Filter to our desired years, take average, and group by taxa and year
coral_cover_7 <- coral_cover_site4 %>%
  filter(Date >= 2009 & Date <= 2017) %>%
  filter(Taxonomy_Substrate_or_Functional_Group == c("Acropora", "Porites", "Pocillopora"))

coral_cover_7$Percent_Cover <- as.numeric(coral_cover_7$Percent_Cover)

coral_cover_mean <- coral_cover_7 %>%
  group_by(Taxonomy_Substrate_or_Functional_Group, Date) %>%
  summarize(mean(Percent_Cover))

acro_coral_cover <- coral_cover_mean %>%
  filter(Taxonomy_Substrate_or_Functional_Group == "Acropora")

acro_coral_cover <- acro_coral_cover[-7,]

kable(acro_coral_cover)
```

### Redefine our time frame for the other datasets

```{r}
# Recruits -----------------------------------------

# Filter for desired years, group by family and year
average_recruits_species_7 <- coral_recruits %>%
  filter(nominal_year >= 2009 & nominal_year <= 2016) %>%
  group_by(nominal_year, family) %>%
  mutate(average_recruits = sum(count))

# Find the average number of recruits each year for each coral species
species_recruits_7 <- average_recruits_species_7 %>%
  group_by( nominal_year, family) %>%
  filter(family == c('Acroporidae')) %>%
  summarize(mean(average_recruits))

# Algae ------------------------------------------

# Filtering to the years we have all data on
algae_7_years <-algae_select %>%
  filter(Year >= 2009 & Year <= 2016)

# Taking the mean nitrogen content by year
annual_average_N_7 <- algae_7_years %>%
  group_by(Year) %>%
  summarize(mean(N))

# Taking the mean algae biomass by year
annual_average_algae_7 <- algae_7_years %>%
 group_by(Year) %>%
  summarize(mean(Dry_Weight))
 
# Temp -------------------------------------------

# Filtering the years of the algae data from 2013-2019
temp_7_years <- yearly_temp_upper %>%
  filter(time_local >= 2009 & time_local < 2017) %>%
  mutate(Year= time_local) %>%
    mutate(Year = year(Year)) %>%
  mutate(Temp = upper_watercolumn_temperature_sbe39_C)%>%
  select(Year, Temp)

```

### Create the Data Frame for the 7 years data

```{r}
acro_data_7 <- data.frame(Year = annual_average_algae_7$Year, 
                        Algae = annual_average_algae_7$`mean(Dry_Weight)`, 
                        Nitrogen = annual_average_N_7$`mean(N)`,
                        Temp = temp_7_years$Temp,
                        Coral_Cover = acro_coral_cover$`mean(Percent_Cover)`,
                        Larval_Recruits = species_recruits_7$`mean(average_recruits)`)
```

```{r}
kable(acro_data_7) %>%
  kable_classic_2(full_width = F)
```

### Filter the data sets by year to join them together


```{r}
# Algae data for each year
algae_2009 <- algae_7_years %>%
  filter(Year == 2009) %>%
  mutate(Temp = 27.21) %>%
  mutate(Recruits = 23) %>%
  mutate(Coral_Cover = 0.22857143)

algae_2010 <- algae_7_years %>%
  filter(Year == 2010) %>%
  mutate(Temp = 27.77) %>%
  mutate(Recruits = 20) %>%
  mutate(Coral_Cover = 0.04883721)

algae_2011  <- algae_7_years %>%
  filter(Year == 2011) %>%
  mutate(Temp = 27.14) %>%
  mutate(Recruits = 107) %>%
  mutate(Coral_Cover = 0.01388889)

algae_2012  <- algae_7_years %>%
  filter(Year == 2012) %>%
  mutate(Temp = 27.58) %>%
  mutate(Recruits = 124) %>%
  mutate(Coral_Cover = 0.03902439)

algae_2013  <- algae_7_years %>%
  filter(Year == 2013) %>%
  mutate(Temp = 27.89) %>%
  mutate(Recruits = 38) %>%
  mutate(Coral_Cover = 0.14285714)

algae_2014  <- algae_7_years %>%
  filter(Year == 2014) %>%
  mutate(Temp = 27.14) %>%
  mutate(Recruits = 31) %>%
  mutate(Coral_Cover = 0.42333333)

algae_2015  <- algae_7_years %>%
  filter(Year == 2015) %>%
  mutate(Temp = 27.80) %>%
  mutate(Recruits = 51) %>%
  mutate(Coral_Cover = 1.76410256)

algae_2016  <- algae_7_years %>%
  filter(Year == 2016) %>%
  mutate(Temp = 27.97) %>%
  mutate(Recruits = 27) %>%
  mutate(Coral_Cover = 0.40625000)


```

### Join the data sets together for our 7 year data frame

```{r}
algae_9_10 <- full_join(algae_2009, algae_2010)

algae_10_11 <- full_join(algae_9_10, algae_2011)

algae_11_12 <- full_join(algae_10_11, algae_2012)

algae_12_13 <- full_join(algae_11_12, algae_2013)

algae_13_14 <- full_join(algae_12_13, algae_2014)

algae_14_15 <- full_join(algae_13_14, algae_2015)

acro_data7 <- full_join(algae_14_15, algae_2016)

kable(summary(acro_data7)) %>%
  kable_classic_2(full_width = F)
```
### Plot the information to see what

```{r}

ggplot(acro_data7, aes(x= Year, y= N, colour = Year)) +
  geom_point() +
ggtitle("Nitrogen Concentration at Moorea LTER Site 4 by Year") +
  xlab("Year") +
  ylab( "Nitrogen Concentration by \n\ Percent Weight of Algal Samples") +
  theme(plot.title = element_text(hjust = 0.5))
```

### Running the linear regression model for the year 7 data

```{r}

acro_total_lm <- lm(Coral_Cover ~ Larval_Recruits + Algae + Nitrogen + Temp + Larval_Recruits:Nitrogen, data= acro_data_7)

summary(acro_total_lm)
```

